---
title: "SICSS: R Bootcamp"
author: "Irina Valenzuela"
date: "6/16/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Bootcamp

## 1. Setting

#### Install Packages 

```{r install-packages, eval=FALSE, warning=FALSE, error=FALSE, message=FALSE}

install.packages('tidyverse') # Data Manipulation, includes ggplot2
```

#### Loading Libraries

```{r Loading libraries, warning=FALSE, error=FALSE, message=FALSE}
library(tidyverse)

```

## 2. Data Wrangling (Cleaning)

#### 2.1 Load data 

Apple mobility data: How far people are going during COVID-19 crisis.

```{r Upload data, warning=FALSE, error=FALSE, message=FALSE}

load("Apple_Mobility_Data.Rdata")
ls() # Show dataframes
head(apple_data,4)
```

#### 2.2 Let's start

Let's start by isolating only the data that corresponds to Brazil


```{r , warning=FALSE, error=FALSE, message=FALSE}

# Apply a filter. 
# Filter on region variable equal Brazil. This subset of data will be a new object called brazil_data

brazil_data <- filter(apple_data, region=="Brazil")
```

Breaking the data set into smaller part. We want the column "region". Then we use the option Select to choose column from a dataframe, and create a new object called regions

```{r , warning=FALSE, error=FALSE, message=FALSE}

regions <- select(apple_data, region)

```



```{r , warning=FALSE, error=FALSE, message=FALSE}

transport_types <- count(apple_data, transportation_type)

# Arrange: sort dataframe. For example arrange/sort by region

alpha_order <- arrange(apple_data, region)

```

Reshaping data
Instead of having many columns for different dates. We want one column with the data, and another with the values. We will use the function gather

```{r , warning=FALSE, error=FALSE, message=FALSE}

# Gather function. New dataset, instead of having lot of columns (date columns), we are going to have lof of rows.  

# key = the name assigned to date variables (dates in our case) , value= new variable we will create. Then, tell the range of variables

# Dates in R: `first date variable`:`last date variable to use`

long_apple_data <- gather(apple_data, key=day, value=mobility_data, `2020-01-13`:`2020-08-20` )

# Now, instead of having 3524 rows and 227 variables, now long_apple_data has 801125 rows and 8 variables (columns)

# Now, we can run simple calculations

```


